\chapter{Litterature review}
\section{Litterature review}
This project and its background has found inspiration by the work presented in [Amundsen, 2025]’s doctoral dissertation "Motion Planning and Control for Underwater Vehicles Operating at Aquaculture Facilities". The literature review is inspired by the one presented in the theses but adapted to fit the scope of this project.
\subsection{Path Planning}
Traditional path planning methods can be divided into graph-based path planning and sampling based path planning. 

\subsubsection{Traditional methods}
Traditional path planning methods can be divided into graph-based path planning and sampling based path planning. Graph based path planning is based on constructing a graph on a formerly known map and using an algortim like Djikstra or A* to find the shortest path between points in the map. Sampling based path planning is based on iteratively expanding the graph to connect two points in the map.Both methods are based on having a predefined static map of the environment. 
\subsubsection{Learning based planning}
papers: 14
In recent years learning based planning has recieved an increase in attention. Papers (insert some papers) describe reinforcement learning methods for path planning. Glasius Bio-Inspired Neural Network
\subsection{Guidance}

\subsubsection{LOS-methods}

\subsubsection{Vecor field guidance}

\subsubsection{pure-pursuit}

\subsection{Autopilot}

\subsubsection{PID}

\subsubsection{sliding mode}

\subsection{Simulation}

\subsection{Net Inspection Solutions} 

\subsubsection{Net pen estimation}


\subsubsection{Net following}